---
title: "BA_HW2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Problem 1:
```{r message=FALSE}
url = "https://raw.githubusercontent.com/jcbonilla/BusinessAnalytics/master/BAData/JC-201709-citibike-tripdata.csv"
citibike = read.csv(url)
df = as.data.frame(citibike)
```


summary statistics for tripduration
```{r message=FALSE}
summary(df$tripduration)
var(df$tripduration)
sd(df$tripduration)
range(df$tripduration)
```


summary statistics for age
```{r message=FALSE}
library(lubridate)
age <- function(dob, age.day = today(), units = "years", floor = TRUE) {
    calc.age = interval(dob, age.day) / duration(num = 1, units = units)
    if (floor) return(as.integer(floor(calc.age)))
    return(calc.age)
}
df$birth.year = as.Date(df$birth.year, "%Y")
df$age = age(df$birth.year)
summary(df$age)
var(df$age, na.rm=TRUE)
sd(df$age, na.rm=TRUE)
range(df$age, na.rm=TRUE)
```


summary statistics for tripduration in minutes
```{r message=FALSE}
df$tripduration_min = round((df$tripduration/60), digits = 2)
summary(df$tripduration_min)
var(df$tripduration_min)
sd(df$tripduration_min)
range(df$tripduration_min)
```


correlation between age and tripduration
```{r message=FALSE}
plot(df$tripduration,df$age)
cor(df$tripduration, df$age, use = "complete.obs")
```


What is the total revenue assuming all users riding bikes from 0 to 45 minutes pay 
3 per ride and user exceeding 45 minutes pay an additional $2 per ride.
```{r message=FALSE}
library(dplyr)
count_less45 = df %>% summarize(count = sum(df$tripduration_min<=45))
count_greater45 = df %>% summarize(count = sum(df$tripduration_min>45))
price_less45 = 3
price_greater45 = 5
total_revenue = (count_less45 * price_less45) + (count_greater45 * price_greater45)
total_revenue[1,1]
```


```{r message=FALSE}
hist(df$tripduration_min,
     xlim=c(0,100), breaks=10000,
     xlab="Min",
     border="blue",
     col="green",
     las=1,
     main="Histogram for Min")
var(df$tripduration_min)
count_less45[1,1]/nrow(df)
```
The variance of trip-duration is huge. The reason is that a small part of clients used citibikes for days
that enlarged the range of data. However, more than 98% of clients' trip durations are less than 45 minutes. This
means although the scope of trip-duration is broad, the significant citibike clients are those people who return the
bikes within 45 min. In other words, the considerable clients of citibike are those people who pay $3 for their rides.


What does this mean for the pricing strategy?

As mentioned above, more than 98% of clients of citibike are those people who pay $3 for their rides. This means
the dominating reason for charging an additional $2 for trips longer than 45min is not generating more revenues. 
Because the revenues generated from the additional fee is only a small portion of total revenue. This pricing stretegy
is more likely to be a signal to the clients that "$3 per ride is a fair price" because most clients would return
the bikes within 45 mins. For citibike, the price strategy of "$3 for all trips from 0 to 45 minutes" can help them 
to generate more revenues compared to "charge by every minute". Under the current price strategy, the average revenue 
per minute is high at the start and gradually goes down. Similarly, the number of clients is also high at the start 
and gradually goes down. So, this pricing strategy maximized revenue because high average revenue corresponds to a 
large amount of clients.


 What does this mean for inventory availability?
 
 An additional 2 is charged to clients with a ride more than 45 mins indicates that citibike expected most clients return 
 the bikes with 45 mins. This indicates the inventory of citybike is not large enough to support too many long trips. By 
 charging an additional $2 for long trips, citibike forces some clients to return the bike instead of keep
 the bikes on their hands for a long period.
 
 
 
Prlblem 2:
```{r message=FALSE}
url2 = "https://raw.githubusercontent.com/jcbonilla/BusinessAnalytics/master/BAData/zagat.CSV"
zagat = read.csv(url2)
df2 = as.data.frame(zagat) 
```

Statistics of Central Tendency
```{r message=FALSE}
sapply(df2[,2:5], mean)
library(psych)
sapply(df2[,2:5], geometric.mean)
sapply(df2[,2:5], harmonic.mean)
sapply(df2[,2:5], median)
library(DescTools)
sapply(df2[,2:5], Mode)
```

Statistics of spread and dispersion of the ratings
```{r message=FALSE}
sapply(df2[,2:5], range)
sapply(df2[,2:5], var)
sapply(df2[,2:5], sd)
sapply(df2[,2:5], IQR)

par(mfrow = c(2,2))
boxplot(df2$Food, xlab="Food", outline = F)
boxplot(df2$Decor, xlab="Decor", outline = F)
boxplot(df2$Service, xlab="Service", outline = F)
boxplot(df2$Price, xlab="Price", outline = F)
```

correlations between rating dimensions
```{r message=FALSE}
source("http://www.sthda.com/upload/rquery_cormat.r")
rquery.cormat(df2[,2:5])
```


weighted average (index) that computes scores for each restaurant. (weighted entropy method) 
```{r message=FALSE}
df2$Name <- NULL
min.max.norm <- function(x){
  (x-min(x))/(max(x)-min(x))
}

w1 = sapply(df2, min.max.norm)

first1 <- function(data)
{
  x <- c(data)
  for(i in 1:length(data))
    x[i] = data[i]/sum(data[])
  return(x)
}

df3 = apply(w1,2,first1)

first2 <- function(data)
{
  x <- c(data)
  for(i in 1:length(data)){
    if(data[i] == 0){
      x[i] = 0
    }else{
      x[i] = data[i] * log(data[i])
    }
  }
  return(x)
}

df4 = apply(df3, 2, first2)

k <- 1/log(length(df4[,1]))
d <- -k * colSums(df4)

d = 1-d

w = d/sum(d)
w
```

What makes a business more profitable?

I believe the first thing to make a business more profitable is to find a balance between
price and the number of customers each day. You do not want too many customers waiting outside, 
and you do not want too many empty tables in the restaurant. The second thing to make a business
more profitable is to make the restaurant unique from the competitors.vMost people would like to 
try something they never met before. To distinguish yourself from others is also essential to 
attract new customers.



If you were hired to advise a new restaurant operator, what would you
recommend in terms of the balance & trade-offs between food, decor, service, and
price?

Based on my previous analysis, the weights of four criteria rank high to low as
Price > Decor > Service > Food. So, the restaurants should put the most effort into managing
their price strategy to attract more customers. Also, they need to put some efforts in
service and decor because these two criteria have similar weights, which are around 20%.